# Reproducible Research: Peer Assessment 1
This is the R markdown document for the assignment.

## Loading and preprocessing the data
First I'm going to require the library and load the data:
```{r}
library(knitr)
data <- read.csv("activity.csv")
head(data)
```


## What is mean total number of steps taken per day?

Since we want to exclude missing values, I am going to first operate on only the data without NA's: 

```{r}
smdata <- na.omit(data)
head(smdata)
```

Next, I will now aggregate the data by date and look at the total number of steps taken each day:

```{r}
sdata <- aggregate( smdata$steps, list(day=smdata$date), sum )
head(sdata)
```

Here is the histogram, the mean, and median of the total number of steps each day:
```{r}
hist(sdata$x, breaks=20, main="Histogram of Total Number of Days", xlab="Total Number of Steps (Summed over Intervals)" )
mean(sdata$x)
median(sdata$x)
```

## What is the average daily activity pattern?
I am assuming that we are still ignore all the NA data here. So I will continue this part of the assignment with smdata that I created above. This time, I will aggregate by interval and average the steps over days:

```{r}
idata <- aggregate( smdata$steps, list(interval=smdata$interval), mean )
head(idata)
tail(idata)
```

As indicated in the instructions, I will now plot this data:

```{r}
plot( idata, type="l" )
title( main="Average Daily Activity Pattern", xlab="Interval", ylab="Number of Steps (Averaged over Days)" )
```

The interval with the maximum number of steps from this data is:

```{r}
which.max( idata$x )
```

Which corresponds with the following interval value:

```{r}
idata$interval[ which.max( idata$x ) ]
```

## Imputing missing values

Since we are looking at the NA's in the data, I am going to switch back to the original data set that was loaded. Here, I am going to check if he interval and date columns have any missing values or not:

```{r}
length( data$date[is.na( data$date ) == TRUE] )
length( data$interval[is.na( data$interval ) == TRUE] )
```

Since neither of them have missing values, then that leaves us with the steps column. Thus, the total number of rows in the data set with NA's is the same as the total number of NA's in the steps column, which can be calculated by the following:

```{r}
length( data$steps[is.na( data$steps ) == TRUE] )
```

This tells us there are 2304 rows with missing values in the data.

To keep things simple, I am just going to replace the missing values using the mean number of steps for the same day. First, I'll create that temporary data frame (without NA's):

```{r}
mdata <- aggregate( smdata$steps, list(day=smdata$date), mean )
```

Then I will create a new data set that is a copy of the original data set:

```{r}
fulldata <- data
head(fulldata)
```

Now I am going to fill in the missing data with the day's mean:

```{r}
# for each row in fulldata
for (i in 1:length(fulldata$steps))
{
  # check if the row has NA in its steps
  if( is.na(fulldata$steps[i]) )
  {
    # convert the date into a Date object
    mydate <- as.Date( fulldata$date[i] )
    
    # see if that date has a mean in mdata
    rez <- mydate == as.Date( mdata$day )
    
    # by default, if mydate does not have a mean,
    # then no steps data exist for that date,
    # so just assign the average to be 0
    myavg <- 0
    
    # otherwise we found the date in mdata,
    # so get that date's mean stored in column x
    if( length( which(rez == TRUE)) > 0 )
    {
      index <- which( rez == TRUE )
      myavg <- mdata$x[index]
    }
    
    # replace NA
    fulldata$steps[i] <- myavg
  }
  # otherwise, the row does not have missing values
}
```

Let's ensure there are 0 NA's left:

```{r}
length(fulldata$steps[is.na(fulldata$steps) == TRUE])
```

Repeating the code from earlier, here is the histogram, the mean, and median of the total number of steps each day found using the imputed data:
```{r}
s2data <- aggregate( fulldata$steps, list(day=fulldata$date), sum )
hist(s2data$x, breaks=20, main="Histogram of Total Number of Days (using Imputed Data)", xlab="Total Number of Steps (Summed over Intervals)" )
mean(s2data$x)
median(s2data$x)
```

Because my imputation strategy replaces missing values with 0 when no mean exists for that date, we see that the histogram displayed this time has a lot more counts for 0. For this reason, the mean is lower than before (it used to be 10766), and the median also got pulled to a smaller number than before (it used to be 10765).

However, there is no impact on the rest of the data when we are looking at the total daily number of steps.

## Are there differences in activity patterns between weekdays and weekends?
