# Reproducible Research: Peer Assessment 1
This is the R markdown document for the assignment.

## Loading and preprocessing the data
First I'm going to require the library and load the data:
```{r}
library(knitr)
data <- read.csv("activity.csv")
head(data)
```


## What is mean total number of steps taken per day?

Since we want to exclude missing values, I am going to first operate on only the data without NA's: 

```{r}
smdata <- na.omit(data)
head(smdata)
```

Next, I will now aggregate the data by date and look at the total number of steps taken each day:

```{r}
sdata <- aggregate( smdata$steps, list(day=smdata$date), sum )
head(sdata)
```

Here is the histogram, the mean, and median of the total number of steps each day:
```{r}
hist(sdata$x, breaks=20, main="Histogram of Total Number of Days", xlab="Total Number of Steps (Summed over Intervals)" )
mean(sdata$x)
median(sdata$x)
```

## What is the average daily activity pattern?
I am assuming that we are still ignore all the NA data here. So I will continue this part of the assignment with smdata that I created above. This time, I will aggregate by interval and average the steps over days:

```{r}
idata <- aggregate( smdata$steps, list(interval=smdata$interval), mean )
head(idata)
tail(idata)
```

As indicated in the instructions, I will now plot this data:

```{r}
plot( idata, type="l" )
title( main="Average Daily Activity Pattern", xlab="Interval", ylab="Number of Steps (Averaged over Days)" )
```

The interval with the maximum number of steps from this data is:

```{r}
which.max( idata$x )
```

Which corresponds with the following interval value:

```{r}
idata$interval[ which.max( idata$x ) ]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
